An Evolutionary Algorithm for the Multiobjective Risk-Equity Constrained Routing Problem

### 2011-11-30 DAW_hazmatHypE/trunk/paper/hazmatRouting_EvolOptim.tex
rewrote conclusions

\subsection{The Feasibility of the EMO Approach}
Due to their heuristic nature, evolutionary multiobjective optimization algorithms like the one presented in this study cannot guarantee that the solutions found are actually Pareto-optimal. However, as we have seen above, the proposed algorithm is able to produce a set of non-dominated solutions in reasonable time which are close to Pareto-optimal solutions found by exact algorithms. Such sets of non-dominated solutions can be a starting point for a decision maker to investigate the trade-offs between the objectives and to decide in which objective space region interesting Pareto-optimal solutions are which can be explored further with more involved algorithms---be it exact algorithms or specialized evolutionary algorithms such as the mentioned W-HypE variant. In the above example instances, it turned out that total risk and cost are indeed more or less fully correlated in the original problem formulation or only loosely (with a larger trade-off between cost and total risk) when the individual arcs' risks and costs are.

Regarding the two initialization schemes, it turned out that in combination with the (local and simple) variation operator, the cost-efficient initialization does not allow the algorithm to find additional solutions besides the initial one and that, in particular if the cost and total risk objectives are correlated, the random initialization is beneficial. Using other initializations as well as additional variation operators from the literature is easily possible with the proposed algorithm but their investigation is left for future research\footnote{For example, the

On the other hand, the proposed variation operator can be changed towards a larger neighborhood by allowing two or more nodes to be added or deleted from each truck path at a time---resulting in more exploration than exploitation when compared to the current variation operator. Also well-designed crossover operators are expected to be beneficial to the problem when, for example, good sub paths can be exchanged between solution paris.
}.

\begin{itemize}

	\item approach is very general and can easily deal with slightly modified problems: e.g. different truck loads (only the objective function code needs to be changed)
	\item it needs to be mentioned that for the tested instances with 311 nodes and 441 arcs, the variation operator of the evolutionary algorithm was able to construct the shortest paths from source to destination in some of the runs for some of the truck paths within the first 100,000 function evaluations. This corresponds to performing random walks of lengths around 20 nodes.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions} \label{S_FW}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The transportation of hazmats is an important optimization problem in the field of sustainable development and in particular the equitable distribution of risks is of high interest. Within this study, we formalize this transportation problem as the minimization of three objectives and propose to use an evolutionary algorithm to cope with the non-linear equity risk objective.

The third objective function of our problem can be rewritten by minimizing the additional variable $z$ as third objective and adding the constraints $\forall q \in Q: z \geq \sum_{c \in C} \sum_{(i,j) \in A} r_{ij}^{cq} f_{ij}^c$. Although this equivalent formulation makes the problem linear (with additional linear constraints), classical algorithms are expected to have difficulties with this formulation as well and our algorithm is supposed to be more efficient in the current formulation due to the fewer number of constraints. Note that, for the moment, the proposed EMO algorithm exists on paper only and an actual implementation has to prove in the future which additional algorithm components (such as problem-specific initialization, recombination operators, or other exact optimization (sub-)procedures) are necessary to generate solutions of sufficient quality and whether adaptively changing the number and capacity of trucks is beneficial.

In future work, one might want to look further into adding (simple) heuristics within the evoluionary algorithm in order to improve the algorithm. For example, one can observe that the solutions produced by the algorithm sometimes uses one and the same arc within a truck path, hence introducing cylces. With the three objectives cost, total and equity risk, such cycles can be easily prevented within the variation operator which might eventually result in a faster convergence of the algorithm to the Pareto front.

% The transportation of hazmats is an important optimization problem in the field of sustainable development and in particular the equitable distribution of risks is of high interest.

% In this study, we formalize transportation of hazmats problem as the minimization of three objectives and propose to use an evolutionary algorithm since many exact optimization procedures have difficulties with the non-linear equity risk objective. Note that our problem formulation can be rewritten by minimizing the additional variable $z$ as third objective and adding the constraints $\forall q \in Q: z \geq \sum_{c \in C} \sum_{(i,j) \in A} r_{ij}^{cq} f_{ij}^c$ \COMMENTD{this has to be adapted according to the new problem formulation}. Although this equivalent formulation makes the problem linear (with additional non-linear constraints), classical algorithms are expected to have difficulties with this formulation as well and the EMO algorithm is supposed to be more efficient in the current formulation due to the fewer number of constraints.
% 
% Note also that, for the moment, the proposed EMO algorithm only exists as a gedankenexperiment and an actual implementation has to prove in the future which additional algorithm components (such as problem-specific initialization, recombination operators, or other exact optimization (sub-)procedures) are necessary to improve the quality of the found solutions to a sufficient level.

>>>>

\subsection{The Feasibility of the EMO Approach}
Due to their heuristic nature, evolutionary multiobjective optimization algorithms like the one presented in this study cannot guarantee that the solutions found are actually Pareto-optimal. However, as we have seen above, the proposed algorithm is able to produce a set of non-dominated solutions in reasonable time. In particular the comparison with the cost-optimal solution shows that the randomized algorithm can find solutions close to the Pareto front. The found sets of non-dominated solutions can be a starting point for a decision maker to investigate the trade-offs between the objectives and to decide in which objective space region interesting Pareto-optimal solutions should be searched for with more involved algorithms---be it exact algorithms or specialized evolutionary algorithms such as the mentioned W-HypE variant.

Regarding the two initialization schemes, it turned out that in combination with the (local and quite simple) variation operator, the cost-optimal initialization does not allow the algorithm to find additional solutions besides the initial one. However, if the cost and total risk objectives are less correlated, this cost-optimal initialization has advantages over the random initialization in the early stage of the search when compared with respect to the hypervolume indicator. In general, however, we suggest to use the random initialization if the number of function evaluations is not a limiting factor.

Using other initializations as well as additional variation operators from the literature is easily possible with the proposed algorithm but their investigation is left for future research. For example, one can think of a ``more random'' initialization where the initial truck paths are generated by iteratively applying the mutation operator on the empty path and also the multiple application of the described mutation operator within one single mutation might be beneficial as the explored space of neighbored paths is larger in comparison to the current mutation. Also well-designed crossover operators are expected to be beneficial to the problem when, for example, good sub paths can be exchanged between solution pairs.

Overall, it is to say that the approach is very general and can easily be adapted if the problem formulationi is (slightly) changed as it is often the case in practice, e.g., if different numbers of trucks are available. Also in case of different truck loads, it is only the objective function code that needs to be changed---a big advantage of blackbox optimizers such as the proposed EMO algorithm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions} \label{S_FW}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The transportation of hazmats is an important optimization problem in the field of sustainable development and in particular the equitable distribution of risks is of high interest. Within this study, we formalize this transportation problem as the minimization of three objectives and propose to use an evolutionary multiobjective optimization (EMO) algorithm to cope with the non-linear equity risk objective.

The algorithm, based on the state-of-the-art environmental selection scheme of HypE, employs a problem-specific variable-length representation and a mutation operator which are known to be well-suited for shortest path problems from theoretical studies. With its random initialization, the algorithm allows to find sets of non-dominated solutions in reasonable time which are close to the Pareto-optimal front. For the tested real-world instances with the road network of the Lazio region and dozens of trucks for up to 4 hazmat commodities, the current Java implementation allows to finish the 100,000 function evaluations of our experiments within less than 15 minutes on a standard Intel Core i5 laptop.

In future work, one might want to look further into adding (simple) heuristics within the EMO algorithm in order to improve the convergence of the algorithm. For example, one can observe that the solutions produced by the algorithm sometimes uses one and the same arc within a truck path, hence introducing cylces. With the three objectives cost, total and equity risk, such cycles can be easily prevented within the variation operator which might eventually result in a faster convergence of the algorithm to the Pareto front.
